Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-10 main.tex
in sec.3 now

We start from a countable set $\VAR=\set{x,y,z,\ldots}$ of variables,
a countable set $\FUN=\set{f,g,\ldots}$ of function symbols, and a
counrable set $\PRED=\set{p,q,\ldots}$ of predicate symbols. Each function symbol and
each predicate symbol has a finite arity. The grammars below generate
the set $\TERM$ of \bfit{terms}, denoted by $s,t,u,\ldots$, the set
$\ATOM$ of \bfit{atoms}, denoted by $a,b,c,\ldots$, and the set
$\FORM$ of \bfit{formulas}, denoted by $A,B,C,\ldots$:
\begin{equation*}
  \begin{array}{r@{~}l}
    t &\coloneqq~ x \mid f(t_1,\dots,t_n)
    \\[.5ex]
    a &\coloneqq~ \ttt\mid\fff\mid p(t_1,\dots,t_n) \mid \dual p(t_1,\dots,t_n)
    \\[.5ex]
    A &\coloneqq~ a \mid A\vlan A \mid A\vlor A \mid \exists x.A \mid \forall x.A
\end{array}
\end{equation*}
where the arity of $t$ and $p$ is $n$. Note, that in this paper we
consider the truth constants $\ttt$ (\emph{true}) and $\fff$
(\emph{false}) as atoms, and we consider all formulas in negation
normalform. The \bfit{negation} $(\dual\cdot)$ is defined for all
atoms and formulas via the De Morgan laws as follows:
\begin{equation*}
  \dual\ttt =\fff
  \quad
  \dual\fff = \ttt
  \quad
  \dual{p(t_1 \ldots t_n)} = \dual{p}(t_1 \ldots t_n)
  \quad
  \dual{\dual p(t_1 \ldots t_n)} = p(t_1 \ldots t_n)
\\
  \dual{\dual a} = a
\end{equation*}

@@

@@

\bfit{Terms} $s, t, u, \ldots$ and \bfit{atoms} (atomic formulas) $\alpha, \beta,
\gamma, \ldots$ are generated inductively from variables $x, y, z, \ldots$: if
$\delta$ (resp. $p$) is an $n$-ary function (resp. predicate) symbol and $t_1, \ldots,
t_n$ are terms, then $\delta t_1 \ldots t_n$ (resp. $pt_1 \ldots t_n$) is a term
(resp. atom). The set of atoms is extended with logical constant $\ttt$ and
$\fff$. We also assume that each predicate symbol $p$ has a \bfit{dual}
predicate symbol $\overline{p}$ with $\overline{p} \neq p$ and
$\overline{\overline{p}} = p$. We extend the duality to atoms with
$\overline{pt_1 \ldots t_n} = \overline{p}t_1 \ldots t_n$, $\overline{\ttt} =
\fff$, and $\overline{\fff} = \ttt$. \bfit{Formulas} $A, B, C$ are generated inductively from atoms by binary
connectives $\cor$ and $\cand$ and quantifiers $\forall x$ and $\exists x$.
The \bfit{negation} $\neg$ is defined in the usual way. A formula is
\bfit{rectified} if all bound variables are distinct from one another and from
all free variables. Without loss of generality, we assume that all formulas are
rectifed throughtout the paper as every formula has a logically equivalent
rectified form.

We consider Gentzen's sequent calculus system $\FOLK$ \cite{gentzen:35:I}.

A \bfit{sequent} is a multiset of formulas and sequents are proved using the
following inference rules:

\begin{figure}[h]
\begin{center}
\begin{prooftree}
  \infer0[$\ax$]{\vdash A, \neg A}
\end{prooftree}
\qquad
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, \neg A}
  \infer2[$\cut$]{\vdash \Gamma, \Delta}
\end{prooftree}
\\[1.5ex]
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \hypo{\vdash \Delta, B}
  \infer2[$\cand$]{\vdash \Gamma, \Delta, A \cand B}
\end{prooftree}
\qquad
\begin{prooftree}
  \hypo{\vdash \Gamma, A, B}
  \infer1[$\cor$]{\vdash \Gamma, A \cor B}
\end{prooftree}
\\[1.5ex]
\begin{prooftree}
  \hypo{\vdash \Gamma, A, A}
  \infer1[$\ctr$]{\vdash \Gamma, A}
\end{prooftree}
\qquad
\begin{prooftree}
  \hypo{\vdash \Gamma}
  \infer1[$\wk$]{\vdash \Gamma, A}
\end{prooftree}
\\[1.5ex]
\begin{prooftree}
\infer0[$\ttt$]{\vdash \ttt}
\end{prooftree}
\qquad
\begin{prooftree}
  \hypo{\vdash \Gamma}
	\infer1[$\fff$]{\vdash \Gamma, \fff}
\end{prooftree}
\\[1.5ex]
\begin{prooftree}
  \hypo{\vdash \Gamma, A}
  \infer1[$\forall$]{\vdash \Gamma, \forall x A}
\end{prooftree}
($x \notin fv(\Gamma)$)
\qquad
\begin{prooftree}
  \hypo{\vdash \Gamma, A[t/x]}
  \infer1[$\exists$]{\vdash \Gamma, \exists x A}
\end{prooftree}
\caption{Gentzen's sequent calculus $\FOLK$}
\end{center}
\end{figure}

>>>>

We start from a countable set $\VAR=\set{x,y,z,\ldots}$ of variables,
a countable set $\FUN=\set{f,g,\ldots}$ of function symbols, and a
counrable set $\PRED=\set{p,q,\ldots}$ of predicate symbols. Each function symbol and
each predicate symbol has a finite arity. The grammars below generate
the set $\TERM$ of \bfit{terms}, denoted by $s,t,u,\ldots$, the set
$\ATOM$ of \bfit{atoms}, denoted by $a,b,c,\ldots$, and the set
$\FORM$ of \bfit{formulas}, denoted by $A,B,C,\ldots$:
\begin{equation*}
  \begin{array}{r@{~}l}
    t &\coloneqq~ x \mid f(t_1,\dots,t_n)
    \\[.5ex]
    a &\coloneqq~ \ttt\mid\fff\mid p(t_1,\dots,t_n) \mid \dual p(t_1,\dots,t_n)
    \\[.5ex]
    A &\coloneqq~ a \mid A\vlan A \mid A\vlor A \mid \exists x.A \mid \forall x.A
\end{array}
\end{equation*}
where the arity of $f$ and $p$ is $n$. Note, that in this paper we
consider the truth constants $\ttt$ (\emph{true}) and $\fff$
(\emph{false}) as atoms, and we consider all formulas in negation
normalform. The \bfit{negation} $(\dual\cdot)$ is defined for all
atoms and formulas via the De Morgan laws as follows:
\begin{equation*}
  \dual\ttt =\fff
  \quad
  \dual\fff = \ttt
  \quad
  \dual{p(t_1 \ldots t_n)} = \dual{p}(t_1 \ldots t_n)
  \quad
  \dual{\dual p(t_1 \ldots t_n)} = p(t_1 \ldots t_n)
\\
  \dual{\dual a} = a
\end{equation*}

@@

In this paper we use the sequent calculus $\FOLK$, shown in
Figure~\ref{fig:LK1}, which is a one-sided variant of Gentzen's
original calculus~\cite{gentzen:35:I} for first-order logic. To
simplify some technicalities later in this paper, we also include here
the $\mixr$-rule. The linear fragment of $\FOLK$, i.e., the fragment
without the rules $\conr$ (\emph{contraction}) and $\weakr$
(\emph{weakening}) defines \emph{first-order multiplicative linear
logic}~\cite{girard:87,girard:88} \emph{with
mix}~\cite{fleury:retore:94,bellin:97}~(MLL1+mix). We denote that system
here with $\FOMLL$ (shown in Figure~\ref{fig:LK1} in the dashed box).

@@
